No need to know: astrophysics-agnostic gravitational wave cosmology

### 2024-03-13 src/tex/ms.tex
[showyourwork] overleaf sync

% Abstract
\begin{abstract}
\Acp{GW} from merging compact objects are the only transient astrophysical sources that encode direct information about their luminosity distances. 
Paired with their redshifts, a Hubble diagram can be constructed to probe the standard cosmological model.  
Although individual signals do not have any reference scale from which a redshift can be inferred, an ensemble of events can provide such a scale via their distribution of source masses. 
This ``spectral siren'' method thus constrains the Universe's expansion rate from \ac{GW} observations alone, without the need for accompanying electromagnetic observations. 
So far, this technique has only been applied with simple parametric representations of the mass distribution. 
However, the use of an incorrect mass model inevitably leads to biases in the cosmological inference, an acute problem given our current lack of understanding of the true astrophysical model.
Furthermore, it is commonly presumed that the exact shape of the mass distribution must be known \emph{a priori} to obtain unbiased measurements of cosmological parameters with the spectral siren methodology.
We demonstrate that unbiased inference of cosmological parameters is possible without any assumptions for the mass distribution's morphology.
We do this by constructing a flexible, Gaussian process-based model for \ac{GW} source population to analyze simulated \ac{GW} data consistent with expectations for the next \acl{LVK} observing run.
We find that both the source mass model and cosmological parameters are correctly reconstructed, and predict a $\variable{output/nonparh0percent.txt}\%$ measurement of \Ho{} and a \result{X\%} measurement of $H(z=\variable{output/mostsensitivez.txt})$ ($1-\sigma$).
This astrophysics-agnostic spectral siren technique will be key to deliver precise and unbiased cosmological constraints with future observations in the presence of possible redshift evolution of the mass spectrum.  
\end{abstract}

>>>>

% Abstract
\begin{abstract}
\Acp{GW} from merging compact objects are the only transient astrophysical sources that encode direct information about their luminosity distances. 
Paired with their redshifts, a Hubble diagram can be constructed to probe the standard cosmological model.  
Although individual signals do not have any reference scale from which a redshift can be inferred, an ensemble of events can provide such a scale via their distribution of source masses. 
This ``spectral siren'' method thus constrains the Universe's expansion rate from \ac{GW} observations alone, without the need for accompanying electromagnetic observations. 
So far, this technique has only been applied with simple parametric representations of the mass distribution. 
However, the use of an incorrect mass model inevitably leads to biases in the cosmological inference, an acute problem given our current lack of understanding of the true astrophysical model.
Furthermore, it is commonly presumed that the exact shape of the mass distribution must be known \emph{a priori} to obtain unbiased measurements of cosmological parameters with the spectral siren methodology.
We demonstrate that unbiased inference of cosmological parameters is possible without any assumptions for the mass distribution's morphology.
We do this by constructing a flexible, Gaussian process-based model for \ac{GW} source population to analyze simulated \ac{GW} data consistent with expectations for the next \acl{LVK} observing run.
We find that both the source mass model and cosmological parameters are correctly reconstructed, and predict a $\variable{output/nonparh0percent.txt}\%$ measurement of \Ho{} and a $\variable{output/nonparhzpercent.txt}\%$  measurement of $H(z=\variable{output/mostsensitivez.txt})$ ($1-\sigma$).
This astrophysics-agnostic spectral siren technique will be key to deliver precise and unbiased cosmological constraints with future observations in the presence of possible redshift evolution of the mass spectrum.  
\end{abstract}
